My title

### 2024-11-29 paper/paper.qmd
Data Cleaning and Variable Creation

>>>>

# 2024
NBA2024 <- read.csv("../data/01-raw_data/All NBA/AllNBA_2024.csv", header = FALSE)
new_col_names <- as.character(unlist(NBA2024[2, ]))
NBA2024 <- NBA2024[-c(1, 2), ]
colnames(NBA2024) <- new_col_names
NBA2024[] <- lapply(NBA2024, type.convert, as.is = TRUE)
```

```{r}

# Years range
years <- 1986:2024

# Normalize columns and calculate MVP index with WS
for (year in years) {
  # Construct dataset name dynamically
  dataset_name <- paste0("NBA", year)

  # Access dataset using `get()`
  dataset <- get(dataset_name)

  # Check if required columns exist
  if (all(c("PTS", "AST", "TRB", "STL", "BLK", "WS") %in% colnames(dataset))) {
    # Calculate column-wise averages, including WS
    averages <- colMeans(dataset[c("PTS", "AST", "TRB", "STL", "BLK", "WS")], na.rm = TRUE)

    # Add average columns to the dataset
    dataset$avg_PTS <- averages["PTS"]
    dataset$avg_AST <- averages["AST"]
    dataset$avg_TRB <- averages["TRB"]
    dataset$avg_STL <- averages["STL"]
    dataset$avg_BLK <- averages["BLK"]
    dataset$avg_WS <- averages["WS"]

    # Normalize by averages
    dataset$norm_PTS <- dataset$PTS / dataset$avg_PTS
    dataset$norm_AST <- dataset$AST / dataset$avg_AST
    dataset$norm_TRB <- dataset$TRB / dataset$avg_TRB
    dataset$norm_STL <- dataset$STL / dataset$avg_STL
    dataset$norm_BLK <- dataset$BLK / dataset$avg_BLK
    dataset$norm_WS <- dataset$WS / dataset$avg_WS

    # Calculate MVP index, now including WS
    dataset$MVP_index <- 0.5 * dataset$norm_PTS +
                         0.2 * dataset$norm_AST +
                         0.2 * dataset$norm_TRB +
                         0.1 * dataset$norm_WS  # Adjust weights as needed

    # Assign the modified dataset back to the environment
    assign(dataset_name, dataset)
  } else {
    stop(paste0("One or more required columns (PTS, AST, TRB, STL, BLK, WS) are missing in dataset for year ", year))
  }
}

# Create a table of the player with the highest MVP index for each year
mvp_table <- data.frame(Year = integer(), Player = character(), MVP_index = numeric(), stringsAsFactors = FALSE)

# Iterate over each year to find the top player
for (year in years) {
  # Construct dataset name dynamically
  dataset_name <- paste0("NBA", year)

  # Access dataset using `get()`
  dataset <- get(dataset_name)

  # Ensure the MVP_index column exists
  if ("MVP_index" %in% colnames(dataset)) {
    # Find the row with the highest MVP_index
    max_row <- dataset[which.max(dataset$MVP_index), ]

    # Extract relevant details (assuming a "Player" column exists)
    mvp_entry <- data.frame(Year = year, 
                            Player = max_row$Player, 
                            MVP_index = max_row$MVP_index,
                            stringsAsFactors = FALSE)

    # Append to the mvp_table
    mvp_table <- rbind(mvp_table, mvp_entry)
  } else {
    stop(paste0("MVP_index column is missing in dataset for year ", year))
  }
}
